---
title: "Homework 3"
subtitle: "Logistic, LDA, QDA, Naive Bayes, and KNN"
format: html
---

```{r include=FALSE}
library(tidyverse)
library(tidymodels)
library(ISLR)
hw_num <- 3
```

## Due 4/23/2024 at 1pm

# Instructions

-   Creating a new *project.* and giving it a sensible name such as `r paste0("homework",hw_num)` and having that project in the course folder you created.

-   Create a new quarto document and give it a sensible name such as `r paste0("hw",hw_num)`.

-   In the `YAML` add the following (add what you don't have). The embed-resources component will make your final rendered `html` self-contained.

```{r}
#| label: yaml_example
#| eval: false
---
title: "Document title"
author: "Your Name"
format:
  html:
    embed-resources: true
---
```
    
-   Though the book used R code for base R, I want you to complete the exercises using functions from *tidymodels*.
-   Set a seed before each problem.
-   Make sure your answers print results. If the output is a large table, use the head() function to shorten the output. 

# Exercises - NOTE THE CHAPTER NUMBER
1.  **Complete exercise chapter 5, number 5 in ISLR.**
 - Replace (b) (iii) with the following: Obtain prediction of default status of each individual in the validation set.

2.  **Complete exercise chapter 4, number 12 in ISLR.**

3.  **Complete exercise chapter 4, number 13 in ISLR.**
NOTE: Confusion matrix is the two-way table. 

4.  **Complete exercise chapter 4, number 16 in ISLR.**

# Submission

When you are finished with your homework, be sure to Render the final document. Once rendered, you can download your file by:

-   Finding the .html file in your File pane (on the bottom right of the screen)
-   Click the check box next to the file
-   Click the blue gear above and then click "Export" to download
-   Submit your final html document to the respective assignment on [Moodle](%7B%7B%3C%20var%20course.moodle%20%3E%7D%7D)
