---
title: "Homework 2"
subtitle: "Cross Validation and tidymodels"
format: html
---

```{r include=FALSE}
library(tidyverse)
library(tidymodels)
library(ISLR)
```

# Instructions

-   Creating a new *project.* and giving it a sensible name such as *homework_2*.

-   Create a new quarto document and give it a sensible name such as *hw2*.

-   In the **YAML** add the following (add what you don't have). The embed-resources component will make your final rendered **html** self-contained.

    ```{r}
    #| label: yaml_example
    #| eval: false
    ---
    title: "Document title"
    author: "my name"
    format:
      html:
        embed-resources: true
    ---
    ```
    
-   Though the book used R code for base R, I want you to complete the exercises using functions from *tidymodels*.
-   Set a seed before each problem.
-   Make sure you answers print results. If the output is a large table, use the head() function to shorten the output. 

# Exercises

1. (ISLR 3) We now review *k*-fold cross-validation.

    a) Explain how *k*-fold cross-validation is implemented.

    b) What are the advantages and disadvantages of *k*-fold cross-validation relative to:

      - The validation set approach?
  
      - LOOCV?

2. (ISLR 8) We will now perform cross-validation on a simulated data set.

    a) Generate a simulated data set as follows below. In this data set, what is $n$ and what is $p$? Write out the model used to generate the data in equation form.

```{r, eval =F}
set.seed(1)
x <- rnorm(100)
y <- x - 2 * x^2 + rnorm(100)
```





2. Continued 

    b) Create a scatterplot of $X$ against $Y$. Comment on what you find.

    c) Set a random seed, and then compute the LOOCV errors that result from fitting the following four models using least squares:

    m1. $Y = \beta_0 + \beta_1 X + \epsilon$
    
    m2. $Y = \beta_0 + \beta_1 X + \beta_2 X^2 + \epsilon$
    
    m3. $Y = \beta_0 + \beta_1 X + \beta_2 X^2 + \beta_3 X^3 + \epsilon$
    
    m4. $Y = \beta_0 + \beta_1 X + \beta_2 X^2 + \beta_3 X^3 + \beta_4 X^4 + \epsilon$

:::{.callout-tip}
You may find it helpful to use the **data.frame()** function to create a single data set containing both $X$ and $Y$.
:::

2. Continued 

    d) Repeat (c) using another random seed, and report your results. Are your results the same as what you got in (c)? Why?

    e) Which of the models in (c) had the smallest LOOCV error? Is this what you expected? Explain your answer.

    f) Comment on the statistical significance of the coefficient estimates that result from fitting each of the models in (c) using least squares. Do these results agree with the conclusions drawn based on the cross-validation results?

3.  The data *FirstYearGPA*, in the *Stat2Data* packages contains measurements on 219 college students. The response variable is GPA (grade point average after one year of college). Use **?FirstYearGPA** to learn about the dataset. 

    a)  Split the data into training and test using 70\% for training and the remaining for test. Fit a multiple linear regression to predict GPA using HSGPA, HU, and White. Give the prediction equation along with the training MSE. Make sure to set a seed. 

    b)  Use the prediction equation in the previous part as a formula to generate predictions of the GPA for each of the cases in the test sample. Then give the test MSE. 

    c)  Now do cross validation instead of using a test and training set. Give the final CV MSE. 

    d)  Which was better - using a test/training approach or using Cross Validation? Why is Cross validation likely to be better in this example? 

    e)    If we had more data, how might we combine both approaches? 

4. The data *Titanic*, in the *Stat2Data* packages a list and outcomes for passengers on the Titanic.

    a)  Fit a logistic regression model using Age and then SexCode using the code below and filling in the blanks. Write down both the logit and probability forms for the fitted model.

```{r}
#| label: log-question
#| eval: false
data(Titanic)
titanic_2 <- as.data.frame(Titanic) |>
  transform(Survived = factor(Survived))

log_spec<- 
  logistic_reg() |>
  set_engine("glm")

logistic_fit <-
  fit(log_spec,
    ____________ ~ ________+________,
    data = ________)

```

4. Continued

    b)  Use the predict function on your model and predict for the data you used to fit the model. 
  
    c)  Using your predicted values, calculate the misclassification rate. You will need to find a way to compare the predicted value to the new estimates and then count when they do not match. 
  
    d)  Now split the data into training and test like in the previous exercise (70/30 split). Give the misclassification rate for the test set. 
  
    e)    Now use cross validation and give the average misclassification rate across folds (**almost** a default output of the collect_metrics function). 

# Submission

When you are finished with your homework, be sure to Render the final document. Once rendered, you can download your file by:

-   Finding the .html file in your File pane (on the bottom right of the screen)
-   Click the check box next to the file
-   Click the blue gear above and then click "Export" to download
-   Submit your final html document to the respective assignment on [Moodle](%7B%7B%3C%20var%20course.moodle%20%3E%7D%7D)
