---
title: "Homework 2"
subtitle: "Cross Validation and tidymodels"
format: html
---

```{r include=FALSE}
library(tidyverse)
library(tidymodels)
library(ISLR)
```

## Due 4/18/2024 at 9am

# Instructions

-   Creating a new *project.* and giving it a sensible name such as *homework_2* and having that project in the course folder you created.

-   Create a new quarto document and give it a sensible name such as *hw2*.

-   In the `YAML` add the following (add what you don't have). The embed-resources component will make your final rendered `html` self-contained.

    ```{r}
    #| label: yaml_example
    #| eval: false
    ---
    title: "Document title"
    author: "my name"
    format:
      html:
        embed-resources: true
    ---
    ```
    
-   Though the book used R code for base R, I want you to complete the exercises using functions from *tidymodels*.
-   Set a seed before each problem.
-   Make sure you answers print results. If the output is a large table, use the head() function to shorten the output. 

# Exercises
1.  **Complete exercise chapter 5, number 3 in ISLR.**

2.  **Complete exercise chapter 5, number 8 in ISLR.**

3.  The data *FirstYearGPA*, in the *Stat2Data* packages contains measurements on 219 college students. The response variable is GPA (grade point average after one year of college). Use `?FirstYearGPA` to learn about the dataset. 

  a)  Split the data into training and test using 70\% for training and the remaining for test. Fit a multiple linear regression to predict GPA using HSGPA, HU, and White. Give the prediction equation along with the training MSE. Make sure to set a seed. 

  b)  Use the prediction equation in the previous part as a formula to generate predictions of the GPA for each of the cases in the test sample. Then give the test MSE. 

  c)  Now do cross validation instead of using a test and training set. Give the final CV MSE. 

  d)  Which was better - using a test/training approach or using Cross Validation? Why is Cross validation likely to be better in this example? 

  e)    If we had more data, how might we combine both approaches? 

4.    The data *Titanic*, in the *Stat2Data* packages a list and outcomes for passengers on the Titanic.

  a)    Fit a logistic regression model using Age and then SexCode using the code below and filling in the blanks. Write down both the logit and probability forms for the fitted model.

```{r}
#| label: log-question
#| eval: false
data(Titanic)
titanic_2 <- as.data.frame(Titanic) |>
  transform(Survived = factor(Survived))

log_spec<- 
  logistic_reg() |>
  set_engine("glm")

logistic_fit <-
  fit(log_spec,
    ____________ ~ ________+________,
    data = ________)

```


  b)    Use the predict function on your model and predict for the data you used to fit the model. 
  
  c)    Using your predicted values, calculate the missclassification rate. You will need to find a way to compare the predicted value to the new estimates and then count when they do not match. 
  
  d)    Now split the data into training and test like in the previous exercise (70/30 split). Give the missclassification rate for the test set. 
  
  e)    Now use cross validation and give the average missclassification rate across folds (`almost` a default output of the collect_metrics function). 

# Submission

When you are finished with your homework, be sure to Render the final document. Once rendered, you can download your file by:

-   Finding the .html file in your File pane (on the bottom right of the screen)
-   Click the check box next to the file
-   Click the blue gear above and then click "Export" to download
-   Submit your final html document to the respective assignment on [Moodle](%7B%7B%3C%20var%20course.moodle%20%3E%7D%7D)
