---
title: "Homework 4"
subtitle: "Regression"
format: html
---

```{r include=FALSE}
library(tidyverse)
library(tidymodels)
library(ISLR)
hw_num <- 3
```

::: {.callout-note collapse="false" appearance="Instructions"}

## Instructions

-   Creating a new *project.* and giving it a sensible name such as `r paste0("homework",hw_num)` and having that project in the course folder you created.

-   Create a new quarto document and give it a sensible name such as `r paste0("hw",hw_num)`.

-   In the `YAML` add the following (add what you don't have). The embed-resources component will make your final rendered `html` self-contained.

```{r}
#| label: yaml_example
#| eval: false
---
title: "Document title"
author: "Your Name"
format:
  html:
    embed-resources: true
---
```
    
-   Though the book used R code from base R, I want you to complete the exercises using functions from *tidymodels* when possible.
-   Set a seed before each problem.
-   Make sure your answers print results. If the output is a large table, use the head() function to shorten the output. 
:::


## Exercises

1.  (ISLR 6.1) We perform **best subset selection**, **forward stepwise selection**, and **backward stepwise selection** on a single data set. For each approach, we obtain $p+1$ models containing $0, 1, 2, \dots, p$ predictors.

a. For a fixed value of $k$, which of the three models with $k$ predictors has the smallest training RSS?

b. For a fixed value of $k$, which of the three models with $k$ predictors has the smallest test RSS?

c. State whether each of the following is **True** or **False**.

   i. The predictors in the $k$-variable model identified by forward stepwise selection are a subset of the predictors in the $(k+1)$-variable model identified by forward stepwise selection.

   ii. The predictors in the $k$-variable model identified by backward stepwise selection are a subset of the predictors in the $(k+1)$-variable model identified by backward stepwise selection.

   iii. The predictors in the $k$-variable model identified by backward stepwise selection are a subset of the predictors in the $(k+1)$-variable model identified by forward stepwise selection.

   iv. The predictors in the $k$-variable model identified by forward stepwise selection are a subset of the predictors in the $(k+1)$-variable model identified by backward stepwise selection.

   v. The predictors in the $k$-variable model identified by best subset selection are a subset of the predictors in the $(k+1)$-variable model identified by best subset selection.


2.  (ISLR 6.8) In this exercise, you will generate simulated data and then use that data to perform model selection.

a. Use the `rnorm()` function to generate a predictor $X$ of length $n = 100$, as well as a noise vector $\epsilon$ of length $n = 100$.

b. Generate a response vector $Y$ of length $n = 100$ according to the model  
$Y = \beta_0 + \beta_1 X + \beta_2 X^2 + \beta_3 X^3 + \epsilon$,  
where $\beta_0$, $\beta_1$, $\beta_2$, and $\beta_3$ are constants of your choice.

c. Use the `regsubsets()` function to perform **best subset selection** in order to choose the best model containing the predictors $X, X^2, \dots, X^{10}$.

   What is the best model according to $C_p$, BIC, and adjusted $R^2$? Show plots that support your answer, and report the coefficients of the selected model.

   Note: You will need to use the `data.frame()` function to create a single data set containing both $X$ and $Y$.

d. Repeat part (c) using **forward stepwise selection** and **backward stepwise selection**. How do your results compare to those from part (c)?

e. Now fit a **lasso** model to the simulated data, again using $X, X^2, \dots, X^{10}$ as predictors.

   Use cross-validation to select the optimal value of $\lambda$. Create plots of the cross-validation error as a function of $\lambda$. Report the resulting coefficient estimates, and discuss the results.

f. Now generate a response vector $Y$ according to the model  
$Y = \beta_0 + \beta_7 X^7 + \epsilon$,  
and perform best subset selection and the lasso. Discuss the results you obtain.






3.  (ISLR 6.9) In this exercise, you will predict the number of applications received using the other variables in the `College` data set.

a. Split the data into a **training set** and a **test set**.

b. Fit a **linear model using least squares** on the training set, and report the **test error** obtained.

c. Fit a **ridge regression** model on the training set, with $\lambda$ chosen by cross-validation. Report the **test error** obtained.

d. Fit a **lasso** model on the training set, with $\lambda$ chosen by cross-validation. Report the **test error** obtained, along with the **number of non-zero coefficient estimates**.

e. Comment on the results obtained. How accurately can we predict the number of college applications received? Is there much difference among the test errors resulting from these three approaches?
  
# Submission

When you are finished with your homework, be sure to Render the final document. Once rendered, you can download your file by:

-   Finding the .html file in your File pane (on the bottom right of the screen).
-   Click the check box next to the file.
-   Click the blue gear above and then click "Export" to download.
-   Submit your final html document to the respective assignment on [Moodle](%7B%7B%3C%20var%20course.moodle%20%3E%7D%7D).
